---
title: "Lab 12 Homework"
author: "Bode W"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(shiny)
library(shinydashboard)
```

## Practice
Let's try to build an app that allows users to explore sexual composition of wolf packs by `pop`. These data are from: Brandell, Ellen E (2021), Serological dataset and R code for: Patterns and processes of pathogen exposure in gray wolves across North America, Dryad, [Dataset](https://doi.org/10.5061/dryad.5hqbzkh51)


1. Start by loading and exploring the data.
```{r}
wolves <- read_csv("data/wolves_dataset.csv")
```

```{r}
names(wolves)
```

2. For the app, here is a faceted version of what we are looking for. We want the reactive part to be `pop` and you should use `shinydashboard`.

```{r}
wolves %>%
  filter(sex != "NA") %>%
  ggplot(aes(x = sex, fill = sex)) +
  geom_bar() +
  facet_wrap(~pop)
```

```{r}
# import libraries
library(tidyverse)
library(shiny)
library(shinydashboard)


pop_options <- unique(wolves$pop)  # get unique values of `pop`

# define UI
ui <- dashboardPage(
    dashboardHeader(title = "Wolves"),
    dashboardSidebar(disable = TRUE),
    dashboardBody(
        fluidRow(
            box(  # this box includes options for user to select
                title = "Options",
                width = 3,
                selectInput(
                    inputId = "pop",
                    label = "Select a `pop`",
                    choices = pop_options
                ),
                selectInput(
                    inputId = "plt_type",
                    label = "Select a plot type",
                    choices = c("bar", "pie")
                )
            ),
            box(  # this box is for rendering the plot
                title = "Plot",
                width = 9,
                plotOutput(outputId = "plot")
            )
        )
    ),
    title = "Wolves"
)

# define server logic
server <- function(input, output, session) {
    plt <- wolves %>%
        select(sex, pop) %>%
        filter(!is.na(sex))

    output$plot <- renderPlot({
        plt <- plt %>%
            filter(pop == input$pop) %>%
            count(sex) %>%
            ggplot()
        switch(
            input$plt_type,
            "bar" = plt + geom_col(aes(x = sex, y = n, fill = sex), alpha = 0.7),
            "pie" = plt +
                geom_col(aes(x = "", y = n, fill = sex), alpha = 0.7) +
                coord_polar("y", start = 0)
        )
    })

    session$onSessionEnded(stopApp)
}

shinyApp(ui, server)
```
