---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.

```{r}
library("tidyverse")
```

```{r}
malaria <- read.csv("./data/malaria.csv")
```

```{r}
head(malaria)
# view(malaria)
```


2. `rdt` refers to rapid diagnostic test and they are identified here by age group.

```{r}
malaria_long <- malaria %>%
    select(-c("malaria_tot")) %>%
    pivot_longer(
        col = starts_with("malaria_rdt_"),
        names_to = "age_group",
        names_prefix = "malaria_rdt_",
        values_to = "cases"
    )
malaria_long # %>% view
```

3. Make the data tidy and store them as a new object.
4. Which district had the highest *total* number of cases on July 30, 2020?

```{r}
malaria_long %>%
    filter(data_date == "2020-07-30") %>%
    group_by(District) %>%
    summarize(total_cases = sum(cases, na.rm = TRUE)) %>%
    arrange(desc(total_cases))
```


```{r}
malaria %>%
    filter(data_date == "2020-07-30") %>%
    group_by(District) %>%
    summarize(
        total_cases = sum(malaria_tot, na.rm = TRUE)
    ) %>%
    arrange(desc(total_cases))
```